---
title: "Homework - Week 5"
author: "Vasco Brazão"
output:
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(rethinking)
library(tidyverse)
library(dagitty)

set.seed(19022022)
```

## 1

```{r one.data}
data("NWOGrants")

d1 <- NWOGrants %>%
  dplyr::mutate(
    gender = dplyr::case_when(
      gender == "m" ~ 1,
      gender == "f" ~ 2
    )
  )
```

Our DAG for this situation:

```{r one.dag}
dag1 <- dagitty::dagitty("dag{
  gender -> discipline;
  gender -> awards;
  discipline -> awards
}")

plot(dag1)
```

To estimate the total effect of gender, we really only need to include gender as a predictor. 

```{r one.model}
m1 <- rethinking::ulam(
  alist(
    awards ~ binomial(applications, p),
    logit(p) <- a[gender],
    a[gender] ~ normal(0, 1)
  ),
  data = d1, chains = 4, cores = 4
)
```

Now to check the model:

```{r one.summary}
precis(m1, depth = 2)
```

This looks OK. In total, it looks like female applicants are at a slight disadvantage, but we best confirm that later by computing the comparison itself. First, some more diagnostics:

```{r one.trace.trank}
traceplot(m1)

trankplot(m1)
```

Wonderful!

To the contrast.

```{r one.contrast}
m1_post <- rethinking::extract.samples(m1)

m1_probs <- tibble(
  p_male = rethinking::inv_logit(m1_post$a[, 1]),
  p_female = rethinking::inv_logit(m1_post$a[, 2]),
  diff_male_female = p_male - p_female
)

m1_probs %>%
  ggplot(aes(x = diff_male_female)) +
  geom_density()
```

It seems for this data set, too, males are at some advantage - around 3 percentage points.

## 2 

```{r two.data}
d2 <- d1 %>%
  mutate(
    discipline = as.integer(discipline)
  )
```

For the direct effect of gender, we'll need to include departments in the analysis and then marginalize over them.

```{r two.model}
m2 <- rethinking::ulam(
  alist(
    awards ~ binomial(applications, p),
    logit(p) <- a[gender, discipline],
    matrix[gender, discipline]:a ~ normal(0, 1)
  ),
  data = d2, chains = 4, cores = 4
)
```

First, a summary:

```{r two.summary}
precis(m2, depth = 3)
```

N_effs and rhats are good. Further diagnostics:

```{r two.trace.trank}
traceplot(m2)

trankplot(m2)
```

Looking pretty good!

Now, to marginalize, following Richard's code. 

```{r two.marge}
# we want to simulate as many apps as there were in the data

n_apps <- sum(d2$applications)

# we also want each department to have as many app as in the data

# apps_per_dept <- sapply(1:6, function (i) sum(d2$applications[d2$discipline==i]))

# can I do it in tidyverse?? i can! is it more pretty though??

apps_per_dept <- purrr::map_dbl(
  .x = 1:9,
  .f = ~ filter(d2, discipline == .x) %>%
    pull(applications) %>%
    sum()
)

# let's get results just for men

p_male <- link(m2, data = list(
  discipline = rep(1:9, times = apps_per_dept),
  applications = rep(1, n_apps),
  gender = rep(1, n_apps)
))

p_female <- link(m2, data = list(
  discipline = rep(1:9, times = apps_per_dept),
  applications = rep(1, n_apps),
  gender = rep(2, n_apps)
))

dens(p_male - p_female)
```

Now it looks like, on average, the direct effect is close to zero, although the distribution is clearly bimodal, with one mode benefitting men and another benefitting women slightly...

## 3

This pattern shows how self-selection (influenced by gender!) onto departments with differing admissions rates can create a picture of gender bias in one direction in the aggregate, all while each department can have their own level of gender bias in any direction.