---
title: "Homework - Week 8"
author: "Vasco Brazão"
output:
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(rethinking)
library(tidybayes)
library(tidyverse)
library(here)


set.seed(02072022)
```

## 1

```{r one.data}
d1_raw <- readr::read_csv(file = here::here("homework", "week-8_vasco", "week08_Monks.csv"))

d1 <- list(
  N_dyads = nrow(d1_raw),
  D = d1_raw$dyad_id,
  like_AB = d1_raw$like_AB,
  like_BA = d1_raw$like_BA
)
```

We adapt the model of the lecture for this.

```{r one.model}
f_1 <- alist(
  like_AB ~ binomial(3, pAB),
  like_BA ~ binomial(3, pBA),
  logit(pAB) <- a + T[D,1],
  logit(pBA) <- a + T[D,2],
  a ~ normal(0, 1),
  
  ## dyad effects
  transpars> matrix[N_dyads, 2]:T <- compose_noncentered(rep_vector(sigma_T, 2), L_Rho_T, Z),
  matrix[2, N_dyads]:Z ~ normal(0, 1),
  cholesky_factor_corr[2]:L_Rho_T ~ lkj_corr_cholesky(2),
  sigma_T ~ exponential(1),
  
  ## corr matrix for dyads
  gq> matrix[2,2]:Rho_T <<- Chol_to_Corr(L_Rho_T) 
)

m1 <- rethinking::ulam(
  f_1, data = d1, chains = 4, cores = 4, iter = 2000
)

precis(m1, depth = 3, pars = c("a", "Rho_T", "sigma_T"))
```

So the correlation of likes within dyad seems to be reliable positive (.45 to .79). However, we haven't taken important features of the data into account yet.

## 2

Now we just add dislike to the model, repeating everything.

```{r two.data}
d2 <- list(
  N_dyads = nrow(d1_raw),
  D = d1_raw$dyad_id,
  like_AB = d1_raw$like_AB,
  like_BA = d1_raw$like_BA,
  dislike_AB = d1_raw$dislike_AB,
  dislike_BA = d1_raw$dislike_BA
)
```

```{r two.model}
f_2 <- alist(
  like_AB ~ binomial(3, pAB),
  like_BA ~ binomial(3, pBA),
  logit(pAB) <- a_like + T[D,1],
  logit(pBA) <- a_like + T[D,2],
  a_like ~ normal(0, 1),
  
  dislike_AB ~ binomial(3, pAB),
  dislike_BA ~ binomial(3, pBA),
  logit(pAB) <- a_dislike + T2[D,1],
  logit(pBA) <- a_dislike + T2[D,2],
  a_dislike ~ normal(0, 1),
  
  ## like dyad effects
  transpars> matrix[N_dyads, 2]:T <- compose_noncentered(rep_vector(sigma_T, 2), L_Rho_T, Z),
  matrix[2, N_dyads]:Z ~ normal(0, 1),
  cholesky_factor_corr[2]:L_Rho_T ~ lkj_corr_cholesky(2),
  sigma_T ~ exponential(1),
  
  ## dislike dyad effects
  transpars> matrix[N_dyads, 2]:T2 <- compose_noncentered(rep_vector(sigma_T2, 2), L_Rho_T2, Z2),
  matrix[2, N_dyads]:Z2 ~ normal(0, 1),
  cholesky_factor_corr[2]:L_Rho_T2 ~ lkj_corr_cholesky(2),
  sigma_T2 ~ exponential(1),
  
  ## corr matrix for dyads
  gq> matrix[2,2]:Rho_T <<- Chol_to_Corr(L_Rho_T), 
  gq> matrix[2,2]:Rho_T2 <<- Chol_to_Corr(L_Rho_T2) 
)

m2 <- rethinking::ulam(
  f_2, data = d2, chains = 4, cores = 4, iter = 2000
)

precis(m2, depth = 3, pars = c("a_like", "a_dislike", "Rho_T", "Rho_T2"))
```

It looks like the correlation is bit lower for disliking. Computing the contrast:

```{r two.contrast}
post2 <- extract.samples(m2)


tidybayes::mean_qi(post2$Rho_T[,1,2] - post2$Rho_T2[,1,2])
```

It does look like we could conclude that there is more reciprocity in liking than disliking, but there is also a lot of uncertainty there.

## 3

```{r three.data}
d3 <- list(
  N_dyads = nrow(d1_raw),
  D = d1_raw$dyad_id,
  like_AB = d1_raw$like_AB,
  like_BA = d1_raw$like_BA,
  dislike_AB = d1_raw$dislike_AB,
  dislike_BA = d1_raw$dislike_BA,
  A = d1_raw$A,
  B = d1_raw$B
)
```

```{r three.model}
f_3 <- alist(
  like_AB ~ binomial(3, pAB),
  like_BA ~ binomial(3, pBA),
  logit(pAB) <- a_like + T[D,1] + R[B,1],
  logit(pBA) <- a_like + T[D,2] + R[A,1],
  a_like ~ normal(0, 1),
  
  dislike_AB ~ binomial(3, pAB),
  dislike_BA ~ binomial(3, pBA),
  logit(pAB) <- a_dislike + T2[D,1] + R[B,2],
  logit(pBA) <- a_dislike + T2[D,2] + R[B,2],
  a_dislike ~ normal(0, 1),
  
  ## like dyad effects
  transpars> matrix[N_dyads, 2]:T <- compose_noncentered(rep_vector(sigma_T, 2), L_Rho_T, Z),
  matrix[2, N_dyads]:Z ~ normal(0, 1),
  cholesky_factor_corr[2]:L_Rho_T ~ lkj_corr_cholesky(2),
  sigma_T ~ exponential(1),
  
  ## dislike dyad effects
  transpars> matrix[N_dyads, 2]:T2 <- compose_noncentered(rep_vector(sigma_T2, 2), L_Rho_T2, Z2),
  matrix[2, N_dyads]:Z2 ~ normal(0, 1),
  cholesky_factor_corr[2]:L_Rho_T2 ~ lkj_corr_cholesky(2),
  sigma_T2 ~ exponential(1),
  
  ## generalized receiving
  transpars> matrix[18,2]:R <- compose_noncentered(sigma_R, L_Rho_R, ZR),
  matrix[2,18]:ZR ~ normal(0, 1),
  cholesky_factor_corr[2]:L_Rho_R ~ lkj_corr_cholesky(2),
  vector[2]:sigma_R ~ exponential(1),
  
  ## corr matrix for dyads
  gq> matrix[2,2]:Rho_T <<- Chol_to_Corr(L_Rho_T), 
  gq> matrix[2,2]:Rho_T2 <<- Chol_to_Corr(L_Rho_T2) ,
  gq> matrix[2,2]:Rho_R <<- Chol_to_Corr(L_Rho_R) 
)

m3 <- rethinking::ulam(
  f_3, data = d3, chains = 4, cores = 4, iter = 2000
)

precis(m3, depth = 3, pars = c("a_like", "a_dislike", "Rho_T", "Rho_T2", "Rho_R"))
```

To show the individual R parameters:

```{r}
individuals <- precis(m3, depth = 3, pars = c("R")) 

ind <- tibble(
  par = rownames(individuals),
  mean = individuals[,1],
  low = individuals[,3],
  high = individuals[,4]
) %>% 
  mutate(
    monk = stringr::str_sub(par, start = 3, end = -4),
    monk = as.numeric(monk),
    type = case_when(
      str_sub(par, start = -2, end = -2) == 1 ~ "like",
      str_sub(par, start = -2, end = -2) == 2 ~ "dislike"
    )
  )

ind %>% ggplot(
  aes(x = monk, color = type, y = mean, ymin = low, ymax = high)) +
    geom_pointrange(position = position_dodge(width = 1)) +
  coord_flip() +
  scale_x_continuous(breaks = seq(0, 18, by = 1))
```

Monks 16 and 10 seem to be particularly disliked.
