---
title: "Week 1 Homework"
author: "Varaidzo Ndebele"
date: "1/13/2022"
output:   
  html_document:
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r libraries, warning = FALSE, message = FALSE}
library(rethinking)
library(tidyverse)
library(ggpubr)
```

## Q1

```{r Q1, warning = FALSE, message = FALSE}

p_grid <- seq( from = 0 , to = 1 , length.out = 1000 )

prior <- rep( 1 , 15 )

likelihood <- dbinom( 4 , size = 15 , prob = p_grid )

unstd.posterior <- likelihood * prior

posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type = "b" ,
      xlab = "probability of water" , ylab = "posterior probability" )

```

## Q2

```{r Q2, warning = FALSE, message = FALSE}

p_grid <- seq( from = 0 , to = 1 , length.out = 1000 )

prior <- ifelse( p_grid < 0.5 , 0 , 1 )

likelihood <- dbinom( 4 , size = 6 , prob = p_grid )

unstd.posterior <- likelihood * prior

posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type = "b" ,
      xlab = "probability of water" , ylab = "posterior probability" )

```

## Q3

```{r Q3, warning = FALSE, message = FALSE}

p_grid <- seq( from = 0 , to = 1 , length.out = 1000 )

prior <- ifelse( p_grid < 0.5 , 0 , 1 )

likelihood <- dbinom( 4 , size = 6 , prob = p_grid )

unstd.posterior <- likelihood * prior

posterior <- unstd.posterior / sum(unstd.posterior)

samples <- sample(p_grid , size = 1e4 , replace=TRUE , prob=posterior )

PI( samples , prob = 0.89 )

HPDI( samples , prob = 0.89 )

```

- The percentile interval (PI) is wider than the highest posterior density interval (HPDI) because the HPDI includes the narrowest region with 89% of the posterior probability (as in the example above), whereas the PI displays the central probability of parameter values, capturing a wider interval (see below)[^1].

- If only the intervals were available, it could be misunderstood that the distribution is more symmetrical than it is.

```{r visualisations, warning = FALSE, message = FALSE}

#visualising plots: the tidyverse way

n <- 1000
n_success <- 4
n_trials  <- 6


  d <- 
    tibble(p_grid = seq(from = 0, to = 1, length.out = n),
       prior) %>% 
  mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(likelihood * prior))


# set no. of samples
n_samples <- 1e4

samples <-
  d %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T)

#visualisation of plots overlayed with defined ontervals
(p1 <- d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() +
  theme_minimal()+
    geom_bracket(
      xmin = 0.52, xmax = 0.88, y.position = 0.0035,
      label = "PI", color = "#1a85ff"
    ) +
  geom_area(data = d %>% filter(p_grid < .88 & p_grid > .52), fill = "#1a85ff", alpha = 0.6) +
  
  geom_line(data = d) +
  geom_line() +
  theme_minimal() +
    geom_bracket(
      xmin = 0.5, xmax = 0.84, y.position = 0.0032,
      label = "HPDI", color = "#d41159"
    ) +
  geom_area(data = d %>% filter(p_grid < .84 & p_grid > .5), fill = "#d41159", alpha = 0.6) +
  labs(title = "HPDI and PI",
    x = "proportion of water (p)",
       y = "density"))
```

[^1]: Adapted from this lovely [bookdown](https://bookdown.org/content/4857/sampling-the-imaginary.html#intervals-of-defined-mass.)
