---
title: "Homework - Week 7"
author: "Vasco Brazão"
output:
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(rethinking)
library(tidybayes)
library(tidyverse)
library(dagitty)


set.seed(24032022)
```

## 1

```{r one.data}
data(bangladesh)

d <- bangladesh

d1 <- d %>% 
  group_by(district) %>% 
  summarise(
    n = n(),
    n_contraception = sum(use.contraception)
  )
```

Let's look at the raw proportions:

```{r one.eda}
p1.1 <- d1 %>% 
  ggplot(aes(x = district, y = n_contraception/n)) +
  geom_point()

p1.1
```

Now we can run a simple varying intercepts model. Let's write it in mathstats first.

$$
C_i \sim \operatorname{Binomial(N_i, p_i)} \\
logit(p_i) = \alpha_{district[i]} \\ 
\alpha_{district} \sim \operatorname{Normal(\bar{\alpha}, \sigma)} \\
\bar{\alpha} \sim \operatorname{Normal(0, 1.5)} \\
\sigma \sim \operatorname{Exponential(1)}
$$

```{r one.model}
m1 <- rethinking::ulam(
  alist(
    n_contraception ~ dbinom(n, p),
    logit(p) <- a[district],
    vector[61]:a ~ dnorm(a_bar, sigma_a),
    a_bar ~ dnorm(0, 1.5),
    sigma_a ~ dexp(1)
  ), data = d1, chains = 4, cores = 4, log_lik = TRUE
)

precis(m1, depth = 2)
```

```{r one.predictions}
posterior <- rethinking::extract.samples(m1)

probs <- rethinking::inv_logit(posterior$a)

means <- apply(probs, 2, mean)

pi <- apply(probs, 2, rethinking::PI)

d_posterior <- tibble::tibble(
  district = 1:61,
  mean = means,
  higher = pi[2,],
  lower = pi[1,],
  highlight = ifelse(district == 54, "yes", "no")
)

p1.1 +
  geom_linerange(aes(y = NULL, ymin = lower, ymax = higher, color = highlight), data = d_posterior, alpha = .5, size = 1.25) +
  geom_point(aes(y = means, color = highlight), data = d_posterior, alpha = .5, size = 2) +
  theme(legend.position = "none")
```

As expected the estimates were shrunk towards the grand mean, much more for those districts with extreme values or little data. District 54, which didn't have any data, still gets an estimate! It is closest to the grand mean and has more uncertainty than the others.

## 2

My DAG:

```{r two.dag}
dag2 <- dagitty::dagitty(
  "dag{
  D -> C;
  D -> U;
  U -> C;
  U -> K;
  K -> C;
  A -> K;
  A -> U;
  A -> D;
  A -> C;
}"
)

plot(dag2)
```

With my DAG, we must do a lot of adjustment. For the direct causal effect, we must adjust by A and D, both common causes of U and C, as well as K, which is in the pipe $U \rightarrow K \rightarrow C$. To get the total causal effect we don't adjust for K.

Let's confirm this analysis:

```{r two.condind}
dagitty::adjustmentSets(
  x = dag2,
  exposure = "U",
  outcome = "C",
  effect = "total"
)

dagitty::adjustmentSets(
  x = dag2,
  exposure = "U",
  outcome = "C",
  effect = "direct"
)
```

Indeed!


## 3

```{r two.data}
d2 <- list(
    D = d$district,
    A = d$age.centered / sd(d$age.centered),
    K = d$living.children,
    U = d$urban,
    C = d$use.contraception,
    alpha = rep(2, 3)
  )
```


For the total causal effect of urban living on contraceptive use, we stratify by age and district. We can adapt the model we used earlier.


```{r two.model.total}
m2 <- rethinking::ulam(
  alist(
    C ~ dbern(p),
    logit(p) <- a[D] + b_age*A + b_urban*U,
    vector[61]:a ~ dnorm(a_bar, sigma_a),
    a_bar ~ dnorm(0, 1.5),
    sigma_a ~ dexp(1),
    b_age ~ dnorm(0, 1.5),
    b_urban ~ dnorm(0, 1.5)
  ), data = d2, chains = 4, cores = 4, log_lik = TRUE
)

precis(m2, depth = 2)
```

According to this model, U is reliably and positively related to contraceptive use. 

However, this is the total effect. What is the direct effect? For that we must also include the number of children in the model. Once again we can practice monotonic effects.

```{r two.model.direct}
m2.1 <- rethinking::ulam(
  alist(
    C ~ dbern(p),
    logit(p) <- a[D] + b_age*A + b_urban*U + b_children*sum(delta_j[1:K]),
    vector[61]:a ~ dnorm(a_bar, sigma_a),
    a_bar ~ dnorm(0, 1.5),
    sigma_a ~ dexp(1),
    b_age ~ dnorm(0, 1.5),
    b_urban ~ dnorm(0, 1.5),
    b_children ~ dnorm(0, 1.5),
    vector[4]: delta_j <<- append_row(0, delta),
    simplex[3]: delta ~ dirichlet(alpha)
  ), data = d2, chains = 4, cores = 4, log_lik = TRUE
)

precis(m2.1, depth = 2)
```

The effect of U is still reliably positive.

I will not be attempting the varying effects at this time.