---
title: "Homework - Week 1"
author: "Vasco Brazão"
date: "Jan 6 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(06012022)

library(tidyverse)
```


## 1

```{r}
grid <- seq(from = 0, to = 1, len = 1000)

prior <- 1

likelihood <- dbinom(x = 4, size = 15, prob = grid)

unstd.posterior <- prior * likelihood

posterior <- unstd.posterior / sum(unstd.posterior)

plot(x = grid, y = posterior)
```

## 2

```{r}
grid <- seq(from = 0, to = 1, len = 1000)

prior <- ifelse(grid < .5, 0, 2)

likelihood <- dbinom(x = 4, size = 6, prob = grid)

unstd.posterior <- prior * likelihood

posterior <- unstd.posterior / sum(unstd.posterior)

plot(x = grid, y = posterior)
```

## 3

```{r}
samples <- sample(
  x = grid,
  size = 1e4,
  replace = TRUE,
  prob = posterior
)

rethinking::PI(samples, prob = .89)

rethinking::HPDI(samples, prob = .89)
```
 
The HPDI is slightly narrower. This is to be expected, since `PI()` chooses the *central* interval with 89% probability, while `HPDI()` finds the *narrowest* interval that contains 89% of the posterior probability.

From the intervals alone, I would likely assume the posterior to be almost symmetrical around them, and would completely miss that no probability at all is assigned to values between 0 and .5. 

## 4

```{r}
samples <- tibble(
  real = sample(
    x = c("W", "L"),
    size = 1e4,
    replace = TRUE,
    prob = c(.7, .3)
  ),
  biased = case_when(
    real == "L" ~ "L",
    real == "W" ~ sample(
      x = c("W", "L"),
      size = 1e4,
      replace = TRUE,
      prob = c(.8, .2)
    )
  )
)

samples %>% count(real, biased) 
```
